\documentclass[11pt]{article}% uses letterpaper by default

%---------- Uncomment one of them ------------------------------
\usepackage[includeheadfoot, top=0.5in, bottom=0.5in, hmargin=1in]{geometry}

% \usepackage[a5paper, landscape, twocolumn, twoside,
%    left=2cm, hmarginratio=2:1, includemp, marginparwidth=43pt,
%    bottom=1cm, foot=.7cm, includefoot, textheight=11cm, heightrounded,
%    columnsep=1cm, dvips,  verbose]{geometry}
%---------------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{graphicx}
\usepackage{varwidth}

\usepackage{amsmath}

% -----------------------------
% Personal config (AT, Sp 2019)
% -----------------------------
\newcommand{\labnumber}{10}  % UPDATE THIS!
\usepackage{enumitem}  % allow one to resume enumerate

% custom figure captioning for floats..
\usepackage[font=small,labelfont=bf]{caption}

\usepackage[colorlinks=true,urlcolor=magenta,linkcolor=blue]{hyperref}
% urlcolor = URL links, linkcolor = within-PDF links

\newcommand*{\mt}{\mathrm}
\newcommand*{\unit}[1]{\;\mathrm{#1}}  % vemod.net/typesetting-units-in-latex
\newcommand*{\abt}{\mathord{\sim}} % tex.stackexchange.com/q/55701
\newcommand*{\Msun}{\mathrm{M}_{\sun}}

% Compact spacing
\setlength\parindent{0pt}
\setlength{\parskip}{0.5em}

\usepackage{datetime2}  % customize \today, defaults to yyyy-mm-dd
\lhead{}
\chead{}
\rhead{}
\renewcommand{\rightmark}{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0.4pt}
\lfoot{\today}  % WARNING - must print day _of_ lab
%\cfoot{\thepage}
% -----------------------------
% End personal config (AT, Sp 2019)
% -----------------------------

\begin{document}

\begin{center}
\Large\textbf{Presentations}
\end{center}

\textbf{When:} Monday, December 6th at 6-9 pm. \\
\textbf{Duration:} $\approx$ 10-13 min, then 0-5 min for
questions/discussion. \\
\textbf{Format:} any combination of slides and/or whiteboard.
Email slides to me before class starts on December 6th. \\
\textbf{Preparation:} 3--6 hours of effort, see next page for details.
Class replaced with open office hour on November 29, where you may come and
discuss as you work.  I'm also available by email or appointment if you
would like to get feedback or do a dry-run of your talk. \\
\textbf{References:} Include references on your slides where appropriate and list them on the last slide in your presentation slides. \\
\textbf{General guideline:} it is often more compelling to discuss a few
sub-topics that you know well, versus putting lots of content that is glossed
over / simply read off your slides.

\vspace{1em}

\textbf{During presentations:} everyone should answer the following questions
for each talk; your responses will be given to the presenter.  Printed forms
will be provided on the day of the presentation.
\setlist{nolistsep}
\begin{itemize}[noitemsep]
    \item What's one thing you learned and/or enjoyed?
    \item What's one strength of the presentation that aided clarity, engagement?
    \item If you were to give the same talk, what would you change to convey
        the ideas more clearly?
\end{itemize}
Come ready to ask questions during and after each talk; these will count for
participation.  Any kind: ``I didn't understand your sentence just now'',
asking for more information, informed disagreement, etc.~will count.

\vspace{1em}

\textbf{Grading:} 18 \% of your final grade (two lab scores).  $8\%$ for the presentations, and rest $8\%$ for your participation (This is excluding the $10\%$ overall participation points)

Content: 75\%
\setlist{nolistsep}
\begin{itemize}[noitemsep]
%\itemsep0em
\item Presenter [ \underline{\hspace{5mm}} ] introduce(s) and describe(s) topic
at level appropriate to this class
\item Presenter [ \underline{\hspace{5mm}} ] explain(s) extent of and
limitations on our knowledge on the topic, including data/observations
underlying knowledge
\item Presenter [ \underline{\hspace{5mm}} ] provide(s) context by drawing
connections to, e.g., different areas of astronomy, concepts from lab or
lecture, other areas of science, areas outside of science, etc.
\item Presenter chooses and cites appropriate references (i.e., goes beyond
Wikipedia and popular press releases).  Presenter submits reference list.
\end{itemize} {\small [ \underline{\hspace{3mm}} ] = easily and concisely (4), sufficiently (3),
is somewhat able to (2), struggles to or cannot (1)
}

Delivery: 25\%
\setlist{nolistsep}
\begin{itemize}[noitemsep]
%\itemsep0em
\item Presentation has a logical flow that audience can [ \underline{\hspace{5mm}} ] follow
\item Presentation aids (slides or boardwork) are [ \underline{\hspace{5mm}} ] understood by audience
\item Presenter speaks clearly, faces audience and makes (or approximates) eye
contact
\item Presenter can [ \underline{\hspace{5mm}} ] address reasonable audience questions
\item Presenter stays within allotted time
\end{itemize}
{\small [ \underline{\hspace{3mm}} ] = easily (4), sufficiently (3), somewhat (2), barely or not (1)}

{\small Adapted from Am.~Astron.~Soc.~Chambliss and other rubrics}

\newpage\pagebreak

\subsection*{Possible topics}

A not-comprehensive list of possible topics are listed below. You can choose something
not listed, so long as it's within the realm of astronomy, our solar system, planets,
exoplanets, and stars.  It should be something you haven't covered in depth in
class or this lab.

I recommend you go one step deeper for most of the below suggestions.
Good topic: ``The Great Red Spot and other storms, vortices, and zonal flows on
Jupiter''.
Not-as-good topic: ``Gas giant atmospheres''.
This will help both you and me determine whether your topic is well-suited for
a 10--13min presentation.

\begin{itemize}[noitemsep]
    \item Solar system planets (including Earth)
        \begin{itemize}[noitemsep]
            \item surface geology and chemistry, prospects for life
            \item atmospheres, climate
            \item magnetospheres, magnetic field, aurora
            \item rings and moons
            \item Earth: formation of the Moon
        \end{itemize}
    \item Smaller solar system bodies
        \begin{itemize}[noitemsep]
            \item Asteroids (incl. Ceres)
            \item Comets (incl. Halley’s Comet)
            \item Kuiper Belt Objects (incl. Pluto)
        \end{itemize}
    \item Sun
        \begin{itemize}[noitemsep]
            \item Interior structure, nuclear fusion, chemistry
            \item Birth, life, and death
            \item Deeper exploration of: sunspots, magnetic reconnection,
                flares and CMEs, tornados
            \item Solar neutrinos
            \item Helioseismology
        \end{itemize}
    \item Planet/star formation
        \begin{itemize}[noitemsep]
            \item Solar system formation and history; meteorites
            \item Age of the solar system
            \item Proto-planetary disks
            \item Planet and planetesimal formation
            \item Brown dwarfs
            \item Star formation (very broad, you'll have to narrow further)
        \end{itemize}
    \item Exoplanets and exo-objects
        \begin{itemize}[noitemsep]
            \item Types of exoplanets: size, mass, composition.  Rocky and
                gaseous planets.
            \item Exoplanet atmospheres, chemistry
            \item Exoplanet detection methods (choose one) and future missions
        \end{itemize}
    \item Spacefaring; Search for Extraterrestrial Intelligence (SETI)
        \begin{itemize}[noitemsep]
            \item Astrobiology, chemistry; the habitable zone
            \item Speciation and extinctions on Earth
            \item Energy usage, Dyson spheres
            \item Communication and signal detection; candidate SETI signals
            \item Space travel; 
        \end{itemize}
    \item Telescopes and spacecraft
        \begin{itemize}[noitemsep]
            \item Ground- versus space-based telescopes
            \item Specific missions/projects: Gemini Planet Imager, Hubble,
                Kepler, TESS (telescopes); Curiosity, OSIRIS-Rex, Dawn, Deep
                Impact, Rosetta/Philae (robots/probes); Juno, Cassini, New
                Horizons, Voyager (orbiters and fly-by spacecraft); many
                others.
            \item NASA budget, mission, proposals.  How funding decisions
                are made.
        \end{itemize}
\end{itemize}

Other options (equally encouraged):
\begin{itemize}
    \item Present a scientific paper.  I recommend looking at the Daily Paper
        Summaries on Astrobites (\url{https://astrobites.org/}).  This is a
        blog that summarizes scientific papers at an introductory level;
        summaries are written by astro graduate students and aimed for
        undergrad/grad students alike.
        Other sources of brief, accessible scientific papers include Nature
        (\url{https://www.nature.com/}), Nature Astronomy
        (\url{https://www.nature.com/natastron/}), and
        Science (\url{https://www.sciencemag.org/}).
        For popular press that can direct you to interesting papers, consider:
        \url{https://www.quantamagazine.org/physics} or
        \url{https://www.scientificamerican.com}.

    \item Biographical study of a famous astronomer or planetary scientist (you can choose your own or from the list below!).
        If you do this, choose at least one scientific contribution to
        emphasize.
        \begin{itemize}[noitemsep]
            \item Galileo Galilei
            \item Johannes Kepler
            \item Caroline Herschel 
            \item Annie Maunder 
            \item Annie Jump Cannon
            \item Cecilia Payne-Gaposchkin 
            \item Carl Sagan
            \item Jill Tarter 
            \item Sara Seager
            \item Vera Rubin
            \item Andrea Mia Ghez 
        \end{itemize}
\end{itemize}

\end{document}
